---
title: "Case_Study_2_2"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Wrangling and Initial Analysis

```{r}
library(dplyr)
library(e1071)
library(plotly)
library(tidyverse)
library(class)
library(caret)

origData = read.csv("/Users/ysojd/Desktop/GRAD SCHOOL COURSEWORK/MSDS_6306_Doing-Data-Science-Master/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/CaseStudy2-data.csv",header = TRUE)

NoSalary <- readxl::read_xlsx("/Users/ysojd/Desktop/GRAD SCHOOL COURSEWORK/MSDS_6306_Doing-Data-Science-Master/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/CaseStudy2CompSet No Salary.xlsx")


NoAttrition <- read.csv("/Users/ysojd/Desktop/GRAD SCHOOL COURSEWORK/MSDS_6306_Doing-Data-Science-Master/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/CaseStudy2CompSet No Attrition.csv",header = TRUE)

```

```{r}
to.be.deleted = which(sapply(origData,function(col) length(unique(col))==1))
origData = origData[,-to.be.deleted]
origData <- mutate_if(origData, is.character, as.factor)
```
##Analysis Plots
```{r}
#Overall Atrition Count
ggplot(data=origData)+
  geom_bar(mapping=aes(x=Attrition))
#Employee attrition count by Department
origData %>% ggplot() + geom_bar(mapping=aes(x=Department, fill=Attrition)) + coord_flip()
#Male vs Female Attrition Count
origData %>% ggplot() + geom_bar(mapping=aes(x=Gender, fill=Attrition)) + coord_flip()
#Satisfaction
origData %>% ggplot()+ geom_bar(aes(x=EnvironmentSatisfaction,fill=Attrition),position="fill")+
  scale_y_continuous(labels = scales::percent)+
  ggtitle("Attrition Due to Environmental Satisfaction")+ylab("Attrition Rate")+xlab("Satisfaction")
# Effects of Job Level, Salary Hike, and Monthly Income on Attrition
pl <- ggplot(origData, aes(x=MonthlyIncome, y=PercentSalaryHike)) + geom_point(shape=2)+ ggtitle("Effect of Job Level(1-5), PercentSalaryHike and MonthlyIncome on Attrition(Y/N)")
pl + facet_grid(Attrition ~ JobLevel)

```

##Create Classification Model
#NB
```{r}
library(lattice)
library(caret)
library(mlbench)
library(rsample)
library(corrplot)
library(h2o)
library(ggplot2)
library(dplyr)
library(tidyr)


control = trainControl(method="repeatedcv", number=10, repeats=3)
#train the model
model = train(Attrition~.,data=origData,method="lvq",preProcess="scale", trControl=control)
#Determine Variable Priority
importance = varImp(model,scale=FALSE)
#summarize importance
print(importance)


origData <- origData %>%
  mutate(
    JobLevel = factor(JobLevel),
    StockOptionLevel = factor(StockOptionLevel),
    TrainingTimesLastYear = factor(TrainingTimesLastYear)
  )

set.seed(13)
split <- initial_split(origData, prop = .7, strata = "Attrition")
train <- training(split)
test  <- testing(split)

# distribution of Attrition rates across train & test set
table(train$Attrition) %>% prop.table()

table(test$Attrition) %>% prop.table()

train %>%
  filter(Attrition == "Yes") %>%
  select_if(is.numeric) %>%
  cor() %>%
  corrplot::corrplot()

train %>% dplyr::select(MonthlyIncome, Education, PerformanceRating, TotalWorkingYears, HourlyRate, JobInvolvement) %>% gather(metric, value) %>% ggplot(aes(value, fill = metric)) + geom_density(show.legend = FALSE) + facet_wrap(~ metric, scales = "free")


features <- setdiff(names(train), "Attrition")
x <- train[, features]
y <- train$Attrition

# set up 10-fold cross validation procedure
train_control <- trainControl(
  method = "repeatedcv", 
  number = 10
  )

# train model
nb.m1 <-suppressWarnings(train(
  x = x,
  y = y,
  method = "nb",
  trControl = train_control
  ))

# results
confusionMatrix(nb.m1)

search_grid <- expand.grid(
  usekernel = c(TRUE, FALSE),
  fL = 0:5,
  adjust = seq(0, 5, by = 1)
)

# train model
nb.m2 <-suppressWarnings(train(
  x = x,
  y = y,
  method = "nb",
  trControl = train_control,
  tuneGrid = search_grid,
  preProc = c("BoxCox", "center", "scale", "pca")
  ))

# top 5 modesl
nb.m2$results %>% 
  top_n(5, wt = Accuracy) %>%
  arrange(desc(Accuracy))
##   usekernel fL adjust  Accuracy     Kappa AccuracySD   KappaSD
## 1      TRUE  1      3 0.8737864 0.4435322 0.02858175 0.1262286
## 2      TRUE  0      2 0.8689320 0.4386202 0.02903618 0.1155707
## 3      TRUE  2      3 0.8689320 0.4750282 0.02830559 0.0970368
## 4      TRUE  2      4 0.8689320 0.4008608 0.02432572 0.1234943
## 5      TRUE  4      5 0.8689320 0.4439767 0.02867321 0.1354681

# plot search grid results
plot(nb.m2)

confusionMatrix(nb.m2)

pred <-suppressWarnings(predict(nb.m2, newdata = test))
confusionMatrix(pred, as.factor(test$Attrition))
```
#Get Average Accuracy, Sensitivity, and Specificity
```{r}
set.seed(13)
iterations = 100
masterAcc = matrix(nrow = iterations,ncol=3)
splitPerc = .75 #Training / Test split Percentage
for(j in 1:iterations)
{
  
  trainIndices = sample(1:dim(origData)[1],round(splitPerc * dim(origData)[1]))
  train = origData[trainIndices,]
  test = origData[-trainIndices,]
  
  model = naiveBayes(Attrition~.,data=train,laplace = 1)
  CM = confusionMatrix(table(predict(model,test),test$Attrition))
  masterAcc[j,1] = CM$overall[1]
  masterAcc[j,2]=CM$byClass[1]
  masterAcc[j,3] = CM$byClass[2]
}

MeanAcc = colMeans(masterAcc)
MeanAcc
```
##Use trained predictive model to measure attrition against NoAttrition dataset.
```{r}
str(NoAttrition)

to.be.deleted2 = which(sapply(NoAttrition,function(col) length(unique(col))==1))
NoAttrition = NoAttrition[,-to.be.deleted2]


NoAttrition <- mutate_if(NoAttrition, is.character, as.factor)
str(NoAttrition)

predNoAttrition = suppressWarnings(predict(nb.m2,NoAttrition))
NoAttrition$Attrition <- predNoAttrition

a = NoAttrition %>% dplyr::select(ID,Attrition)

write.csv(a, "/Users/ysojd/Desktop/GRAD SCHOOL COURSEWORK/MSDS_6306_Doing-Data-Science-Master/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/Case2PredictionsYvanSojdehei Attrition.csv", row.names = FALSE)
```
##Analysis and creation of Monthly Income Regession Model
```{r}
histogram.curve <- hist(origData$MonthlyIncome, breaks = 10, col = "purple", xlab = "Monthly Income", main = "Histogram with Normal Curve")
# Adding normal curve to the histogram
xfit <- seq(min(origData[,19]), max(origData[,19]), length=40)
yfit <- dnorm(xfit, mean=mean(origData[,19]), sd=sd((origData[,19])))
yfit <- yfit*diff(histogram.curve$mids[1:2])*length(origData$MonthlyIncome)
lines(xfit, yfit, col ="black", lwd=2)
```

#Create Regression Model
```{r}
library(MASS)

set.seed(24)
train.control <- trainControl(method = "cv", number = 10)
reg.model = train(MonthlyIncome~., data=origData,
                   method="lmStepAIC",
                   trControl = train.control,
                   trace=FALSE)
reg.model$results
reg.model$finalModel
```
```{r}
summary(reg.model$finalModel)
```
#Utilize Model to predict monthly incomes for No Salary Case Study Data.
```{r}
to.be.deleted = which(sapply(NoSalary,function(col) length(unique(col))==1))
NoSalary = NoSalary[,-to.be.deleted]
NoSalary <- mutate_if(NoSalary, is.character, as.factor)
NoSalary <- mutate_if(NoSalary, is.double, as.numeric)
NoSalary$JobLevel <- as.factor(NoSalary$JobLevel)
NoSalary$StockOptionLevel <- as.factor(NoSalary$StockOptionLevel)
NoSalary$TrainingTimesLastYear <- as.factor(NoSalary$TrainingTimesLastYear)

MonthlyIncome = predict(reg.model,NoSalary)
MonthlyIncome

NoSalary$MonthlyIncome <- MonthlyIncome
b = NoSalary%>% dplyr::select(ID,MonthlyIncome)

write.csv(b,"/Users/ysojd/Desktop/GRAD SCHOOL COURSEWORK/MSDS_6306_Doing-Data-Science-Master/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/Case2PredictionsYvanSojdehei Salary.csv" ,row.names = FALSE)
```
#Additional Analysis
```{r}
library(GGally)
library(dplyr)
origData = read.csv("/Users/ysojd/Desktop/GRAD SCHOOL COURSEWORK/MSDS_6306_Doing-Data-Science-Master/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/CaseStudy2-data.csv",header = TRUE)

set.seed(1)
IncomePerDegree<-data.frame(origData$MonthlyIncome/origData$Education)
#normalize <- function(x) {
#  return ((x - min(x)) / (max(x) - min(x))) }
#IncomePerDegree <- as.data.frame(lapply(IncomePerDegree, normalize))
IncomePerDegree <- as.data.frame(IncomePerDegree)
IncomePerDegree <- setNames(IncomePerDegree, "Incomeperdegree")
origData<-cbind(origData,IncomePerDegree)

# Effect of Gender, Education and Income Per Level of Education on Attrition(Y/N)
pl <- ggplot(origData, aes(x=origData$Incomeperdegree, y=Education, color=EducationField)) + geom_point(size=3, shape=2)+ ggtitle("Effect of Gender, Education and Income Per Level of Education on Attrition(Y/N)") + xlab("Monthly Income By Degree")
pl + facet_grid(Attrition ~ Gender)
#Education Level as it relates to Income
origData %>% dplyr::select(Incomeperdegree,Education,Attrition,JobRole) %>%
  ggplot(aes(x=Education, y=Incomeperdegree, color=JobRole)) +
  geom_point(size=3, shape=20, position="jitter") +
  geom_smooth(method=lm) +
  ggtitle("Education Level vs Income")+xlab("Education Level")+ylab("Income Per Degree")

#Look at Gender as it relates to Attrition
origData$Attrition <- as.factor(origData$Attrition)
origData$Gender <- as.factor(origData$Gender)
origData %>% 
  ggplot(aes(x=Gender, y=Attrition ,fill=Department)) +
  geom_point(size=2, shape=23,  position="jitter") +
  geom_smooth(method=lm) +
  ggtitle("Gender vs. Attrition")+xlab("Gender")+ylab("Attrition")

```



























